---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Homework 1 - Report
Alfieri Anna and Bejarano Montalvo Ana Lucy
11 april, 2019

##Ex. 0
As a starting point you can install the packages we’ll extensively use throughout the semester:
* readr
* tidyverse

```{r}
install.packages("readr")
install.packages("tidyverse")
library(readr)
library(tidyverse)
```

##Exercise 1
Try to figure out the answers without executing the code. Check your answers in R Studio.

a. Given the vector: x <- c("ww", "ee", "ff", "uu", "kk"), what will be the output for x[c(2,3)] ?

The code x[c(2,3)] should give as an output the values of the vector x for the second and third row, namely "ee" and "ff". The following code also confirms that.
```{r}
x <- c("ww", "ee", "ff", "uu", "kk")
x[c(2,3)]
```

b. Let a <- c(2, 4, 6, 8) and b <- c(TRUE, FALSE, TRUE, FALSE), what will be the output for the R expression max(a[b])?

The code max(a[b]) looks for the maximum values of a given b. This means that it will search for the highest value in vector a that corresponds to true in vector b. The result should be 6. We applied the following code to check our assumption.
```{r}
a <- c(2, 4, 6, 8)
b <- c(TRUE, FALSE, TRUE, FALSE)
max(a[b])
```


c. Is it possible to apply the function my_function_name using x and a as arguments?



##Exercise 2
Consider a vector x such that: x=c(1, 3, 4, 7, 11, 18, 29) Write an R statement that will return a list X2 with components of value: x*2, x/2, sqrt(x) and names ‘x*2’, ‘x/2’, ‘sqrt(x)’.
```{r}
x<-c(1, 3, 4, 7, 11, 18, 29)
X2<-list("x*2"=x*2,"x/2"=x/2, "sqrt(x)"=sqrt(x))
X2
```

##Exercise 3
Read the file Table0.txt into an object DS.
```{r}
Table0 <- read_table2("Table0.txt",col_names =FALSE, 
                      cols(X1 = col_character(),X2 = col_double(), X3 = col_double(), X4 = col_double(),
                           X5 = col_character())
)
View(Table0)
```
1. What is the data type for the object DS?
It's a table.

2. Change the names of the columns to Name, Age, Height, Weight and Sex.
```{r}
Table0<-setNames(Table0, c("Name", "Age", "Height", "Weight", "Sex"))
```

3. Change the row names so that they are the same as Name, and remove the variable Name.
```{r}

```

4. Get the number of rows and columns of the data.

##Exercise 4
Convert DS from the previous exercise to a data frame DF.
```{r}
Table0<-as.data.frame(Table0)
```
Add an additional column “zeros” in DF with all elements 0.
```{r}

```

Remove the Weight comuln from DF.
