---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
#### Exercise 3: House Prices

In this exercise, you will work with the [house prices dataset obtained from Kaggle](https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data).
The dataset contains 81 columns describing (almost) every aspect of residential homes in Ames, Iowa.

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=6,
  echo =TRUE,#Show codes
  cache = FALSE,
  messages = FALSE,#whether to preserve messages. if FALSE, all messages will be printed in the console instead of the output document. For example error and descriptions
  results = 'hide',
  global.par = TRUE, fig.align = 'center', out.width='0.7\\linewidth') #Wide of plot and center
```


```{r message=FALSE, warning=FALSE}
library("readr")
library("dplyr")
library("ggplot2")
library("kableExtra")
```


__3.a__ Save the [data file](https://raw.githubusercontent.com/tvatter/dsfba_2019/homework/data/house_prices.csv) in `data` folder of your project. Load the dataset into the global environment by using function `read_csv` and assign it to variable `house_prices`. Make sure that the class of `house_prices` is a tibble.

```{r, message = FALSE}
house_prices<-read_csv("data/house_prices.csv")
class(house_prices)
```

__3.b__ You will work with only four variables, namely, `LotArea`, `KitchenQual`, `LotShape`, and `SalePrice`, which indicate area (in square feet), kitchen quality, general shape, and sale price (in dollars) of property. 
Modify `house_prices` to have only these four columns, as well as transform `LotArea` from square feet into square meters. Bonus: try to use only one `dplyr` function.

```{r, message = FALSE}
house_prices<-house_prices%>%
  transmute(LotArea_m=(LotArea/10.76391), KitchenQual, LotShape, SalePrice)
```

__3.c__ For how many lots the sale price was greater than its mean value?

```{r, message = FALSE}
house_prices%>%
  filter(SalePrice>mean(SalePrice))%>%
  summarise(n_lots_high_price=n())
```
The sale price of 560 lots were greater than its mean value.

__3.d__ Display the average sale price for each kitchen quality level.

  Note, levels of `KitchenQual` correspond to the followin values:

    - Ex -- Excellent,
    - Gd -- Good,
    - TA -- Typical/Average,
    - Fa -- Fair
    - Po -- Poor

```{r, message = FALSE}
house_prices%>%
  group_by(KitchenQual)%>%
  summarise(Average_sale_price= mean(SalePrice))%>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = F)
```

__3.e__ Display ten randomly selected observations from the dataset. What happens if you execute your code a few times? How can you make sure that each execution returns the exact 10 rows?

```{r, message = FALSE}
house_prices%>%
  sample_n(size=10)%>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```
If the code is executed a few times can be noticed that the rows selected change.

```{r}
set.seed(32)
house_prices%>%
  sample_n(size=10)%>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```
To be able to reproduce the same results it is possible to set a seed. The number of seed specify a vector of random numbers that is going to be used. 

__3.f__ Print out the minimum price of observations for which shape is regular (i.e., `"Reg"`), and the kitchen has excellent quality.

```{r, message = FALSE}
house_prices%>%
  filter(LotShape=="Reg" & KitchenQual=="Ex")%>%
  summarise(Min_price=min(SalePrice))
```

__3.g__ Produce boxplots with kitchen quality as the x-axis and the price as the y-axis. Regroup kitchen quality in the following order: excellent, good, typical/average, and fair. What can you say about the relationship between kitchen quality and the price?

```{r, message = FALSE}
house_prices%>%
  ggplot(mapping=aes(x=KitchenQual, y=SalePrice))+
  ggtitle("Relationship betweeen kitchen quality and price")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_boxplot()
```
```{r}
house_prices%>%
  ggplot(mapping=aes(x=KitchenQual, y=SalePrice)) +
  geom_boxplot(mapping=aes(x=factor(KitchenQual, level= c("Ex", "Gd", "TA", "Fa"))))+
  ggtitle("Relationship betweeen kitchen quality and price")+
  theme(plot.title = element_text(hjust = 0.5))
```
The price of a residential home in Ames, Iowa decreases whith the level quality of the kitchen.

__3.h__ Draw a scatter chart to investigate the dependence between `LotArea` and `SalePrice`. Further, use different colors depending on the kitchen quality, and different shapes depending on the shape of the property. 

```{r, message = FALSE}
house_prices%>%
  ggplot(mapping = aes(x=LotArea_m, y=SalePrice))+
  ggtitle("Relationship betweeen sales price and lot area")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_point()
```
```{r}
house_prices%>%
  ggplot(mapping = aes(x=LotArea_m, y=SalePrice, color=KitchenQual, shape=LotShape))+
  ggtitle("Relationship betweeen sales price and lot area")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_point()+
  scale_shape_manual(values=c(0, 3, 16,17))+
  scale_color_manual(values=c("red", "green", "orange","bisque3"))
```

